Road Accident Prediction in Time and Space

### 2019-04-19 paper/paper.tex
Change positive and negative examples generation part of paper

\begin{abstract}
Road accidents are an important issue of our modern societies, responsible for millions of deaths and injuries. In this paper we show how one can leverage open datasets from a City like Montreal, Canada, to create some accident prediction models, using state-of-the-art big data analytics methods in Python. Such models could then be used in the context of road accident prevention, but also to identify key factors that can lead to a road accident, and eventually, help to elaborate new policies. This study also explains how we dealt with the severe class imbalance issue of the accident prediction problem, a recurrent issue that touches many other scientific domains like medical diagnoses or fraud detection. In particular, we show how we implemented Balanced Random Forests, a variant of the Random Forests machine learning algorithm, into the Scala and Python APIs of Apache Spark big data framework. All source code can be found on Github.
\end{abstract}

@@

\section{Introduction}
\subsection{Context}
According to the World report on road traffic injury prevention, 2004, more than one million people die every year from car accidents and dozens of millions of people are being injured\cite{Peden2004}. In this report, the World Health Organization also describe road traffic systems as “the most complex and the most dangerous system with which people have to deal every day”. This issue is all the more important that the number of accidents is forecast to continue rising. In the last decade, Big Data Analytics has been emerging \cite{Gandomi2015}, allowing scientists to handle a very large amount of data, but also complex and heterogeneous data, in order to extract useful insights from it. In the context of accident prediction, Big Data Analytics could provide insights on the conditions leading to an increased risk of road accidents that could then be used to develop traffic-related policies and prevention operations. Our contributions include: (1) a demonstration of how simple and easily available datasets can be assembled to obtain meaningful features for road accident prediction, (2) a comparison of three different machine learning algorithm dealing with data imbalance in the context of road accident prediction, (3) the implementation of Balanced Random Forest introduced by Chen et al. in Apache Spark for efficient distributed training. Finally, all the source code used is publicly available on Github under the MIT license.

\subsection{Related work}
Accident prediction has been extensively studied in the last decade. The main techniques that have been used at the beginning of the research in road accidents prediction were methods using the Poisson distribution and the negative binomial regression \cite{Chang2005}. Then, as machine learning methods become more popular, algorithms like decision trees have been used in several forms like classification and regression decision trees, frequent pattern trees and random forests
\cite{Chang2005, Theofilatos2017, Abellan2013, Lin2015}, along with other machine learning algorithms like Support Vector Machines \cite{LI20081611}. Recently, the outstanding progress of deep learning algorithms have resulted in their extensive use in every field, including in Geospatial machine learning (Convolutional Neural Network and Long Short-Term Memory neural network for example) \cite{Yuan2018, QChen2016}.
\section{Methods}
\subsection{Choosing a Big Data framework: Apache Spark or Dask}
According to Dask’s official website, “Dask is a flexible library for parallel computing in Python” \cite{dask}. Apache Spark, on the other hand, describes itself as “a fast and general-purpose cluster computing system” \cite{spark}. Both of them are accepted tools with wide supporting community, even though Apache Spark is a bigger and older project. Although we almost exclusively used Apache Spark’s Python API (called “pySpark”) in this study, we started our experiment using Dask. Some advantages of Dask over Spark made us chose it for our analysis in the first place; For example, Dask “is developed in coordination with […] projects like Numpy, Pandas and Scikit-Learn” \cite{dask}, all widely used Python packages in machine-learning and data science. It is also admitted that Dask is lighter than Spark and designed to be more flexible in terms of applications and algorithms. However, we found ourselves writing a lot more code using Dask than Spark for the same result. Spark map-reduce-based interface is very handy and allow easier processing of datasets. For all these reasons we found out that Spark fitted our needs best and we switched to it at an early time of the study.

@@

\subsection{Datasets}
We used three public datasets provided by the city of Montreal and the government of Canada: (1) Montreal Vehicle Collision: This dataset provided by the city of Montreal contains all the road collisions reported by the police in Montreal from 2012 to 2018. For each accident, the dataset contains the date and localization of the accident, information on the number of injuries and death, the number of vehicles involved, and information on the road condition. We used only the localization and the date of the accident since we do not have the other piece of information when no accident happened. Another dataset with all vehicle collision in Canada is available but without the localization of the accident, therefore we restrained our analysis to the city of Montreal. (2) National Road Network: This dataset provided by the government of Canada contains the geometry of all roads in Canada. For each road segment, a few meta-data are given. Sadly, for roads in Quebec the speed limit and the surface type are not provided. The data was available in various format, we chose to use the Keyhole Markup Language, which is easy to read because it is based on the Extensible Markup Language (XML). (3) Data Historical Climate Data: This dataset provided by the government of Canada contains hourly weather information measured at different weather stations across Canada. For each station and every hour, the dataset provides the temperature, the humidity, the wind direction and speed, the visibility, the atmospheric pressure and observations of atmospheric phenomenon such as snow, fog, rain, etc.

\subsection{Positive and negative samples generation}
Our machine learning problem can be stated as a binary classification problem, the positive class being the occurrence of an accident and the negative class being all the cases where an accident did not happen. We got a little bit less than 150 000 positive samples from the national collision dataset and we generated a random sub-sample of 5 million negative samples (over 10 billion possibilities). The generation of the negative samples basically consisted in a Cartesian product between the roads from the road network dataset and all the hours of all days that occurred between the 1st January 2012 and the 31 of December 2017, the dates range of the Montreal Vehicle Collision dataset. This huge cartesian product operation together with the joins with the data from the other datasets made our dataset generation pretty heavy. It resulted in an impractical amount of time and memory space requirements to generate the dataset. Our first implementation was querying the Historical Climate Data API in real-time with a cache mechanism. Our idea was to collect only the weather stations and hours necessary for our sample of negative examples, but it resulted in bad performances. We got a performance increase by first building a Spark data frame with all the Historical Climate Data for weather stations around Montreal and then joining the two datasets. We conducted a detailed analysis of our algorithm to improve its performances. We notably obtained a good performance increase by not persisting intermediary results of the road segment identification for accidents. As opposed to what we initially thought, recomputing these results was faster than writing and reading them in the cache. Finally, the identification of the road segment corresponding to accidents was very memory intensive, we modified this step to be executed by batch of one month.  With these improvements and a few other tricks including partitioning the data frame at key points in our algorithm, we manage to reduce the processing time from about one minute to generate about 1000 positive samples to less than 1 second. We also used a cluster from “Compute Canada” to take maximum advantage of Apache Spark distributed nature for the generation of examples and the hyper-parameter tuning of our models.

@@

For the features from the road segments, we were limited by the limited metadata provided on the road segments. From the shape of the road segment, we computed the length of the road segment, and from the name of the street we identified the type of road (Highway, street, boulevard, etc.). In addition road segment are classified in three different level in the dataset depending on their importance in the road network, we created a categorical feature from this information. For these two categorical features, we encoded them as suggested in Element of Statistical Learning\cite{elementsofstat} in section 9.2.4, instead of using one hot encoding which would create an exponential number of splits, we index the categorical variable ordered by the proportion of the example belonging to the given category which are positive samples. This encoding guarantee to provide optimal splits on these categorical variables. Lastly, we added an a feature giving the number of accidents which occurred previously on this road segment.

@@

 Our source code is available on Github at \href{https://github.com/GTimothee/accident-prediction-montreal}{https://github.com/GTimothee/accident-prediction-montreal}.

>>>>

\begin{abstract}
Road accidents are an important issue of our modern societies, responsible for millions of deaths and injuries. In this paper we show how one can leverage open datasets from a city like Montreal, Canada, to create some accident prediction models, using state-of-the-art big data analytics methods in Python. Such models could then be used in the context of road accident prevention, but also to identify key factors that can lead to a road accident, and eventually, help to elaborate new policies. This study also explains how we dealt with the severe class imbalance issue of the accident prediction problem, a recurrent issue that touches many other scientific domains like medical diagnoses or fraud detection. In particular, we show how we implemented Balanced Random Forests, a variant of the Random Forests machine learning algorithm, into the Scala and Python APIs of Apache Spark big data framework. All source code can be found on Github.

\end{abstract}

@@

\section{Introduction}
\subsection{Context}
According to the World report on road traffic injury prevention, 2004, more than one million people die every year from car accidents and dozens of millions of people are being injured\cite{Peden2004}. In this report, the World Health Organization also describe road traffic systems as “the most complex and the most dangerous system with which people have to deal every day”. This issue is all the more important that the number of accidents is forecast to continue rising. In the last decade, Big Data Analytics has been emerging \cite{Gandomi2015}, allowing scientists to handle a very large amount of data, but also complex and heterogeneous data, in order to extract useful insights from it. In the context of accident prediction, Big Data Analytics could provide insights on the conditions leading to an increased risk of road accidents that could then be used to develop traffic-related policies and prevention operations. Our contributions include: 
\begin{itemize}
\item a demonstration of how simple and easily available datasets can be assembled to obtain meaningful features for road accident prediction,
\item a comparison of three different machine learning algorithm dealing with data imbalance in the context of road accident prediction,
\item the implementation of Balanced Random Forest introduced by Chen et al. in Apache Spark for efficient distributed training. 
\end{itemize}
All the source code used is publicly available on Github under the MIT license.\footnote{https://github.com/GTimothee/accident-prediction-montreal}

\subsection{Related work}
Accident prediction has been extensively studied in the last decade. The main techniques that have been used at the beginning of the research in road accidents prediction were methods using the Poisson distribution and the negative binomial regression \cite{Chang2005}. Then, as machine learning methods become more popular, algorithms like decision trees have been used in several forms like classification and regression decision trees, frequent pattern trees and random forests
\cite{Chang2005, Theofilatos2017, Abellan2013, Lin2015}, along with other machine learning algorithms like Support Vector Machines \cite{LI20081611}. Recently, the outstanding progress of deep learning algorithms has resulted in their extensive use in every field, including in Geospatial machine learning (Convolutional Neural Network and Long Short-Term Memory neural network for example) \cite{Yuan2018, QChen2016}.
\section{Methods}
\subsection{Choosing a Big Data framework: Apache Spark or Dask}
According to Dask’s official website, “Dask is a flexible library for parallel computing in Python” \cite{dask}. Apache Spark, on the other hand, describes itself as “a fast and general-purpose cluster computing system” \cite{spark}. Both of them are accepted tools with wide supporting community, even though Apache Spark is a bigger and older project. Although we almost exclusively used Apache Spark’s Python API (called “pySpark”) in this study, we started our experiment using Dask. Some advantages of Dask over Spark made us chose it for our analysis in the first place; For example, Dask “is developed in coordination with […] projects like Numpy, Pandas and Scikit-Learn” \cite{dask}, all widely used Python packages in machine-learning and data science. It is also admitted that Dask is lighter than Spark and designed to be more flexible in terms of applications and algorithms. However, we found ourselves writing a lot more code using Dask than Spark for the same result. Spark map-reduce-based interface is very handy and allow easier processing of datasets. For all these reasons we found out that Spark fitted our needs best and we switched to it at an early time of the study.

@@

\subsection{Datasets}
We used three public datasets provided by the city of Montreal and the government of Canada: 
\begin{itemize}
\item Montreal Vehicle Collision: This dataset provided by the city of Montreal contains all the road collisions reported by the police in Montreal from 2012 to 2018. For each accident, the dataset contains the date and localization of the accident, information on the number of injuries and death, the number of vehicles involved, and information on the road condition. We used only the localization and the date of the accident since we do not have the other piece of information when no accident happened. Another dataset with all vehicle collision in Canada is available but without the localization of the accident, therefore we restrained our analysis to the city of Montreal.
\item National Road Network: This dataset provided by the government of Canada contains the geometry of all roads in Canada. For each road segment, a few meta-data are given. Sadly, for roads in Quebec the speed limit and the surface type are not provided. The data was available in various format, we chose to use the Keyhole Markup Language, which is easy to read because it is based on the Extensible Markup Language (XML). 
\item Data Historical Climate Data: This dataset provided by the government of Canada contains hourly weather information measured at different weather stations across Canada. For each station and every hour, the dataset provides the temperature, the humidity, the wind direction and speed, the visibility, the atmospheric pressure and observations of atmospheric phenomenon such as snow, fog, rain, etc.
\end{itemize}

\subsection{Positive and negative examples generation}
Our machine learning problem can be stated as a binary classification problem, the positive class being the occurrence of an accident and the negative class being the non-occurence of an accident on a given road at a given date and hour. The vehicle colission dataset contains 150 thousands of collisions, among them $134 489$ contain the date, the hour and the localisation of the accident. For each of these accidents, we identify the closest road segment. These pairs of time and  road segment are used as positive examples. For the negative examples, we generated a random sample of 5 millions combinations of time and road segment which did not appear in the collision dataset.  The generation of these examples together with the join operation with the data from the other datasets made our dataset generation pretty heavy. It resulted in an impractical amount of time and memory space requirements to generate the dataset. Our first implementation was querying the Historical Climate Data API in real-time with a cache mechanism. Our idea was to collect only the weather stations and hours necessary for our sample of negative examples, but it resulted in bad performances. We got a performance increase by first building a Spark data frame with all the Historical Climate Data for weather stations around Montreal and then joining the two datasets. We conducted a detailed analysis of our algorithm to improve its performances. We notably obtained a good performance increase by not persisting intermediary results of the road segment identification for accidents. As opposed to what we initially thought, recomputing these results was faster than writing and reading them in the cache. Finally, the identification of the road segment corresponding to accidents was very memory intensive, we modified this step to be executed by batch of one month.  With these improvements and a few other tricks including partitioning the data frame at key points in our algorithm, we managed to reduce the processing time to a reasomable time. We also used a cluster from “Compute Canada” to take maximum advantage of Apache Spark distributed nature for the generation of examples and the hyper-parameter tuning of our models.

@@

For the features from the road segments, we were limited by the limited metadata provided on the road segments. From the shape of the road segment, we computed the length of the road segment, and from the name of the street we identified the type of road (Highway, street, boulevard, etc.). In addition road segment are classified in three different level in the dataset depending on their importance in the road network, we created a categorical feature from this information. For these two categorical features, we encoded them as suggested in The Elements of Statistical Learning\cite{elementsofstat} in section 9.2.4, instead of using one hot encoding which would create an exponential number of splits, we index the categorical variable ordered by the proportion of the example belonging to the given category which are positive samples. This encoding guarantee to provide optimal splits on these categorical variables. Lastly, we added an a feature giving the number of accidents which occurred previously on this road segment.

@@
